---
title: "Install Tensorflow"
author: "Daniel Morillo"
output: html_notebook
---

# Introduction

This notebook installs the environment components needed to interact with
TensorFlow from R. It follows [the "Tensorflow for R" installation
guide](https://tensorflow.rstudio.com/install/).

# Installation

```{r config}
# The repository URL needs to be specified when calling `install.packages()`
#   from a non-interactive session:
REPO_URL <- "https://cran.rstudio.com"
```

We will not be using GPUs, but in order to make it simple, we will use the
standard installation steps. Therefore, we will just follow the normal steps.
First, we install the TensorFlow package for R:

```{r install-tensorflow}
# Install TensorFlow package:
install.packages("tensorflow", repos = REPO_URL)
```

When creating the virtual environment after installing the Python environment,
there seems to be a problem when R is restarted after installing Keras, making
the virtual environment in the new Python installation to not be used (this may
be due to having a prior version of Python installed). So instead of installing
a version of Python automatically, I download manually the latest version from
[the download page](https://www.python.org/downloads/) and install it manually
(in Windows the [direct download
link](https://www.python.org/ftp/python/3.11.3/python-3.11.3-amd64.exe) can be
used):

```{r config-python}
# Install and configure Python installation:
library(reticulate)

path_to_python <- install_python()

virtualenv_create("r-reticulate", python = path_to_python)
```

The tutorial gives us two options at this step: either install TensorFlow
directly, or install Keras, which includes TensorFlow and some other common
packages. After trying the second option (in case I needed any of those packages
later on), I install the TensorFlow library directly:

```{r install-keras}
# Install Tensorflow;
library(tensorflow)
install_tensorflow()
```

# Verification

I verify that the installation of TensorFlow is correct:

```{r check-installation}
# Confirm installation of TensorFlow:
library(tensorflow)
tf$constant("Hello Tensorflow!")
```

## Installation of virtual environment

In a previous attempt, the installation of Tensorflow seemed to fail because a
virtual environment was missing. If that happens, run the following (and then
run again chunk `check-installation` above):

```{r install-miniconda-env}
reticulate::install_miniconda()
```

# Versioning

As a last step, we make sure to add all the new packages to the R environment:

```{r version-renv}
renv::snapshot(prompt = FALSE)
```
